---
title: "API Practice Day 3 EDS 214"
output: html_document
date: "2022-08-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Hands-on
USGS dataRetrieval R package to retrieve hydrological data.USGS is managing a vast network of gauges to monitor freshwater across the US.

Exercise 1
Create a new repository on GitHub named api-practice

Clone is to Taylor

```{r}
library(dataRetrieval)
library(ggplot2)
library(tidyverse)
```

# Read in data 
Webpage: https://waterdata.usgs.gov/nwis/uv?site_no=11118500
```{r}
# Ventura River Site
siteNumber <- "11118500" 
ChoptankInfo <- readNWISsite(siteNumber)
parameterCd <- "00060"

#Raw daily data:
rawDailyData <- readNWISdv(siteNumber,parameterCd,
                      "2019-10-01","2020-10-05")
```



# Start a new Markdown document to plot the discharge time-series for the Ventura River from 2019-10-01 to 2020-10-05

```{r}
ggplot(data = rawDailyData, 
       aes(x = Date,
           y = X_00060_00003),
       color = X_00060_00003) +
  geom_line(color = "orange") +
  theme_minimal() +
  labs(title = "Ventura River 2019 - 2020",
       x = "Date",
       y = "Discharge")
  
  
```


Tutorial for the package can be found here: https://cran.r-project.org/web/packages/dataRetrieval/vignettes/dataRetrieval.html#daily-data

Bonus
How would you try to determine when this stream gauge record started using the API?

metajam
The metajam R package relies on the dataONE API to download data and metadata into your R Environment. It is currently supporting KNB, ADC and EDI repositories because they rely on the metadata standard EML.

Short intro to the package

Exercise 2
Let’s determine what percentage of Alaskan household are speaking only English!

The data: https://doi.org/10.5063/F1N58JPP
# Download the data in 
Download the data household_language.csv using metajam
Read the data into R using metajam
```{r}
data_obj <- "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3A7fc6f6db-c5ea-426a-a743-1f2edafb43b8"
path <- "~/R/eds214_api_practice"
# download data and metadata
library(metajam)
download_d1_data(data_obj, path)
```


Read the metadata
```{r}
read_d1_files("~/github/eds214_api_practice2/doi_10.5063_F1CJ8BPH_household_language_csv")

dataframe <- data$data

```


Write a piece of code that will compute the percentage of Alaskan household speaking only English for the year 2009 to 2015
Create a plot to visualize this data
Bonus
How does it compare to French?